Comparison of Catalogs

### 2024-09-30 compare/compare.tex
completeness notes+rm percent_hist

Completeness checks are crucial for addressing selection biases,
ensuring accurate redshift distributions, and validating galaxy
simulations. They help identify gaps in the data and guide follow-up
observations, ensuring that the catalog provides a reliable sample for
scientific analysis. Without these checks, conclusions drawn from the
data may be inaccurate or incomplete.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

Checking for completeness in galaxy catalogs is essential to ensure that
the data accurately represents the true population of galaxies.
Incomplete catalogs can lead to biased results in statistical studies,
such as the distribution of galaxy luminosity, mass, or star formation
rates. Additionally, missing galaxies, especially those at faint
magnitudes or large distances, can distort cosmological measurements and
hinder our understanding of galaxy formation and evolution. Completeness
checks are crucial for addressing selection biases, ensuring accurate
redshift distributions, and validating galaxy simulations.

Distance-based corrections are applied to mitigate these biases by
adjusting for the underrepresentation of galaxies at greater distances.
As galaxies move farther away, they become fainter and harder to detect,
leading to a drop in the number of detected galaxies. Methods like
\textbf{volume corrections} (e.g., V/Vmax) and \textbf{luminosity
function-based corrections} help account for these effects by estimating
the true galaxy population based on the observed sample. These
corrections ensure that statistical analyses, even in incomplete
catalogs, more accurately reflect the full galaxy population.

@@

\subsection{Completeness of the Catalogs, based on the Distance and the
Morphological
Type}\label{completeness-of-the-catalogs-based-on-the-distance-and-the-morphological-type}

@@

\begin{table}

\caption{\label{tbl-dis-comp}Table showing the Completeness of the
Catalogs, per bin}

\subcaption{\label{tbl-dis-comp-1}HECATE}

\begin{tabular}{rrll}
\toprule
Bin Start & Bin End & Unique \% & Inner \%\\
\midrule
0 & 0 & 43\% & 57\%\\
0 & 1 & 93\% & 7\%\\
1 & 1 & 95\% & 5\%\\
1 & 2 & 98\% & 2\%\\
2 & 2 & 98\% & 2\%\\
2 & 2 & 95\% & 5\%\\
2 & 3 & 98\% & 2\%\\
3 & 3 & 96\% & 4\%\\
3 & 4 & 96\% & 4\%\\
4 & 4 & 95\% & 5\%\\
4 & 4 & 96\% & 4\%\\
4 & 5 & 90\% & 10\%\\
5 & 5 & 81\% & 19\%\\
5 & 6 & 86\% & 14\%\\
6 & 6 & 62\% & 38\%\\
6 & 6 & 58\% & 42\%\\
6 & 7 & 74\% & 26\%\\
7 & 7 & 71\% & 29\%\\
7 & 7 & 62\% & 38\%\\
7 & 8 & 64\% & 36\%\\
8 & 8 & 52\% & 48\%\\
8 & 9 & 71\% & 29\%\\
9 & 9 & 69\% & 31\%\\
9 & 9 & 77\% & 23\%\\
9 & 10 & 75\% & 25\%\\
10 & 10 & 73\% & 27\%\\
10 & 11 & 73\% & 27\%\\
11 & 11 & 83\% & 17\%\\
\bottomrule
\end{tabular}

\subcaption{\label{tbl-dis-comp-2}LCV}

\begin{tabular}{rrll}
\toprule
Bin Start & Bin End & Unique \% & Inner \%\\
\midrule
0 & 1 & 68\% & 32\%\\
1 & 2 & 50\% & 50\%\\
2 & 3 & 67\% & 33\%\\
3 & 4 & 80\% & 20\%\\
4 & 5 & 85\% & 15\%\\
5 & 6 & 79\% & 21\%\\
6 & 6 & 73\% & 27\%\\
6 & 7 & 83\% & 17\%\\
7 & 8 & 74\% & 26\%\\
8 & 9 & 72\% & 28\%\\
9 & 10 & 85\% & 15\%\\
10 & 11 & 78\% & 22\%\\
\bottomrule
\end{tabular}

\end{table}%

@@

As we can see from the histograms Figure~\ref{fig-dis-comp} and
Figure~\ref{fig-type-comp} the sample of nique galaxies of each catalog,
gets smaller by an almost constant proportion (Inner join).

@@

@@

\includegraphics{compare_files/figure-pdf/cell-16-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-18-output-1.pdf}

\includegraphics{compare_files/figure-pdf/cell-20-output-1.pdf}

@@

@@

\textbf{Percentage change:}

\includegraphics{compare_files/figure-pdf/cell-24-output-1.pdf}

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
& diff\_T & diff\_T\_clip \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
count & 229 & 191 \\
mean & -30 & -1 \\
std & 119 & 10 \\
min & -1000 & -39 \\
25\% & -11 & -2 \\
50\% & 0 & 0 \\
75\% & 0 & 2 \\
max & 131 & 30 \\
\end{longtable}

{[}?{]} After the sigma clip we only lose 39 galaxies (\(14\%\)) and we
can see that both the median and the mean of the percentage change are
close to \(0\%\).This is why we can assume that the Types of the
galaxies are the same for the two catalogs

\paragraph{Normalize the scale of galaxy
types}\label{normalize-the-scale-of-galaxy-types}

It is very possible that the two catalogs use different scaling methods,
as indicated by the use of decimal numbers in HECATE.

\begin{longtable}[]{@{}lllll@{}}
\toprule\noalign{}
& T & TType & T\_norm & TType\_norm \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
count & 229 & 287 & 229 & 287 \\
mean & 7 & 7 & 1 & 1 \\
std & 5 & 5 & 0 & 0 \\
min & -5 & -3 & 0 & 0 \\
25\% & 7 & 6 & 1 & 1 \\
50\% & 10 & 9 & 1 & 1 \\
75\% & 10 & 10 & 1 & 1 \\
max & 10 & 10 & 1 & 1 \\
\end{longtable}

Also, as we can see the minimum values are lower by 2 in HECATE, which
complies with the linear fit.

\includegraphics{compare_files/figure-pdf/cell-27-output-1.pdf}

\includegraphics{compare_files/figure-pdf/cell-28-output-1.pdf}

\includegraphics{compare_files/figure-pdf/cell-29-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-30-output-2.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-31-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-33-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-35-output-1.pdf}

@@

@@

\includegraphics{compare_files/figure-pdf/cell-40-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-42-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-44-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-46-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-49-output-1.pdf}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-karachentsevSTARFORMATIONPROPERTIES2013a}
Karachentsev, Igor D., and Elena I. Kaisina. 2013. {``{STAR FORMATION
PROPERTIES IN THE LOCAL VOLUME GALAXIES VIA Hα AND FAR-ULTRAVIOLET
FLUXES}.''} \emph{AJ} 146 (3): 46.
\url{https://doi.org/10.1088/0004-6256/146/3/46}.

\bibitem[\citeproctext]{ref-karachentsevUPDATEDNEARBYGALAXY2013}
Karachentsev, Igor D., Dmitry I. Makarov, and Elena I. Kaisina. 2013.
{``{UPDATED NEARBY GALAXY CATALOG}.''} \emph{AJ} 145 (4): 101.
\url{https://doi.org/10.1088/0004-6256/145/4/101}.

@@

>>>>

Completeness checks are crucial for addressing selection biases,
identify gaps in the data and guide follow-up observations, ensuring
that the catalog provides a reliable sample for scientific analysis.
Without these checks, conclusions drawn from the data may be inaccurate
or incomplete.

\subsection{Completeness of the LCV
Catalog}\label{completeness-of-the-lcv-catalog}

The local volume selection has been made by taking into account galaxies
with: 1. Radial velocities of
\begin{equation}\phantomsection\label{eq-lcv-vel}{
    V_{LG}<600 \text{ km}\cdot\text{m}^{-1}
    }\end{equation} 2. Distances of D\textless11 Mpc

A simultaneous fulfillment of both conditions (1) and (2) is not
required.

\begin{itemize}
\tightlist
\item
  Completeness within a 10 Mpc radius is difficult to assess due to:

  \begin{itemize}
  \tightlist
  \item
    Variability in galaxy properties (luminosity, size, surface
    brightness, gas content)
  \item
    Errors in distance measurements (Tully--Fisher method errors of
    \textasciitilde20--25\%), especially at the 10 Mpc boundary.

    \begin{itemize}
    \tightlist
    \item
      Accurate distances are mainly known within \textasciitilde5 Mpc.
    \end{itemize}
  \item
    Non-Hubble motions (\textasciitilde300 km/s) may make up half of the
    adopted velocity constraint Equation~\ref{eq-lcv-vel}

    \begin{itemize}
    \tightlist
    \item
      Solution for our usage: only keep the galaxies inside a radius of
      \(D=11 Mpc\)
    \item
      HI sources in surveys with low angular resolution: ``The presence
      around our Galaxy of hundreds of high-velocity clouds with low
      line-of-sight velocities and small W50 widths also provokes the
      inclusion of false ``nearby'' dwarf galaxies in the
      LV''Karachentsev, Makarov, and Kaisina (2013)
    \item
      photographic emulsion defect -\textgreater{} exotic cases of
      galaxies -\textgreater{} radial velocity of \(+614 km s^{−1}\)
    \end{itemize}
  \end{itemize}
\item
  Astro-Spam and Survey Errors:

  \begin{itemize}
  \tightlist
  \item
    Automatic surveys produce false detections (e.g., stars
    misclassified as galaxies, high-velocity clouds mistaken for
    dwarfs).
  \end{itemize}
\item
  Conditional Completeness Estimate:

  \begin{itemize}
  \tightlist
  \item
    Galaxies brighter than \(M_B^c = -11^m\) or with linear diameters
    larger than \(A_{26} = 1.0\) kpc show (40--60)\% completeness.
  \item
    ``among the members of the Local Group (D \textless{} 1 Mpc), only
    half of the galaxies have absolute magnitudes brighter than −11m.
    Consequently, more than half of the ultra-faint dwarf companions
    around normal galaxies, like the Sombrero galaxy (D = 9.3 Mpc),
    still remain outside our field of view.''
  \end{itemize}
\item
  Undetected Ultra-Faint Dwarfs:

  \begin{itemize}
  \tightlist
  \item
    Many faint dwarf galaxies remain undetected beyond the Local Group.
    Estimated population of undetected dwarfs could be as large as
    \(10^3–10^4\) within the LV.
  \end{itemize}
\item
  Surface Brightness Distribution:

  \begin{itemize}
  \tightlist
  \item
    Surface brightness remains consistent across distances, except for
    ultra-faint dwarfs (\(SB <31 \text{mag}\cdot\text{arcsec}^{-2}\)).
  \item
    Faintest dwarf galaxies are detectable only nearby due to their
    resolution into individual stars.
  \end{itemize}
\item
  Luminosity-Size Relationship:

  \begin{itemize}
  \tightlist
  \item
    Observations align with cosmological models predicting \(L\sim A^3\)
    Navarro, Frenk, and White (1996), though deviations occur for
    extremely low surface brightness galaxies.
  \item
    ``The deviation from it at the extremely low surface brightness end
    is due to a systematic overestimation of dwarf galaxy sizes, the
    brightness profiles of which lie entirely below the Holmberg
    isophote.''
  \end{itemize}
\end{itemize}

\subsection{Completeness of HECATE}\label{completeness-of-hecate}

The completeness of HECATE is difficult to assess due to: - Unknown
selection function of HyperLEDA and selection effects from other
cross-correlated catalogs. - Estimation based on comparing B-band
luminosity distribution with the galaxy luminosity function (LF).

\begin{itemize}
\item
  HECATE is:

  \begin{itemize}
  \item
    Complete down to \(L_B\sim 10^{9.5} L_{B,\odot}\) for distances less
    than 33 Mpc.
  \item
    Complete down to \(L_B\sim 10^{10} L_{B,\odot}\) for distances
    between 67 Mpc and 100 Mpc.
  \item
    Incomplete at distances greater than 167 Mpc, even for the brightest
    galaxies.
  \end{itemize}
\item
  Completeness estimates based on B-band luminosity density:

  \begin{itemize}
  \item
    \textgreater75\% complete for distances less than 100 Mpc.
    \(\sim 50\%\) complete at distances of \(\sim170\) Mpc.
  \item
    Completeness exceeds 100\% within 30 Mpc due to the overdensity of
    galaxies around the Milky Way.
  \end{itemize}
\item
  Completeness in terms of stellar mass (M*):

  \begin{itemize}
  \tightlist
  \item
    Similar to B-band completeness when measured with K\_s-band
    luminosity as a tracer for stellar mass. Overdensity at small
    distances and cut-off at large distances are observed.
  \end{itemize}
\item
  Completeness in terms of star formation rate (SFR):

  \begin{itemize}
  \item
    \textasciitilde50\% complete between 30 and 150 Mpc.
  \item
    Lower SFR completeness due to limitations in WISE-based SFR
    estimates, which lack full sky coverage.
  \item
    Despite IRAS's limited depth, it provides \textgreater50\% coverage
    for star-forming galaxies in the local neighborhood.
  \end{itemize}
\end{itemize}

In this section we will check the completeness

@@

\subsection{Completeness of the Data}\label{completeness-of-the-data}

\subsubsection{Distance}\label{distance}

@@

@@

As we can see from the histograms Figure~\ref{fig-dis-comp} and
Figure~\ref{fig-type-comp} the sample of unique galaxies of each
catalog, gets smaller by an almost constant proportion (Inner join).

@@

\includegraphics{compare_files/figure-pdf/cell-13-output-1.pdf}

@@

@@

\includegraphics{compare_files/figure-pdf/cell-17-output-1.pdf}

\includegraphics{compare_files/figure-pdf/cell-19-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-22-output-1.pdf}

``Morphological type of galaxy in the numerical code according to the
classification by de Vaucouleurs et al.~(1991). It should be noted that
about three quarters of objects in the LV are dwarf galaxies, which
require a more detailed morphological classification. For example, dwarf
spheroidal galaxies and normal ellipticals are usually denoted by the
same numerical code T \textless{} 0, although their physical properties
drastically differ. The classification problem arises as well for the
``transient'' type dwarf galaxies, T r, which combine the features of
spheroidal (Sph) and irregular (Ir)systems. Due to small classification
errors, such objects may ``jump'' from one end of the T scale to the
other.'' Karachentsev, Makarov, and Kaisina (2013)

\begin{itemize}
\tightlist
\item
  we can assume that the galaxies from the upper left and lower right
  regions of the plot are classified differently because of this
\item
  this might explain the large errors E\_T of HECATE
\item
  why we see a difference in the distribution of
  Figure~\ref{fig-type-comp}
\end{itemize}

We can remove the ``problematic'' galaxies by only keeping the ones
with: \[
|{T_{HECATE}-T_{LCV}}|<7
\]

@@

@@

\begin{minipage}{0.50\linewidth}
\includegraphics{compare_files/figure-pdf/cell-29-output-1.pdf}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics{compare_files/figure-pdf/cell-29-output-2.pdf}\end{minipage}%
\newline
\begin{minipage}{\linewidth}

@@

@@

\includegraphics{compare_files/figure-pdf/cell-32-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-34-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-38-output-1.pdf}

@@

@@

\includegraphics{compare_files/figure-pdf/cell-41-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-43-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-45-output-1.pdf}

@@

\includegraphics{compare_files/figure-pdf/cell-48-output-1.pdf}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-karachentsevSTARFORMATIONPROPERTIES2013a}
Karachentsev, Igor D., and Elena I. Kaisina. 2013. {``{STAR FORMATION
PROPERTIES IN THE LOCAL VOLUME GALAXIES VIA Hα AND FAR-ULTRAVIOLET
FLUXES}.''} \emph{The Astronomical Journal} 146 (3): 46.
\url{https://doi.org/10.1088/0004-6256/146/3/46}.

\bibitem[\citeproctext]{ref-karachentsevUPDATEDNEARBYGALAXY2013}
Karachentsev, Igor D., Dmitry I. Makarov, and Elena I. Kaisina. 2013.
{``{UPDATED NEARBY GALAXY CATALOG}.''} \emph{The Astronomical Journal}
145 (4): 101. \url{https://doi.org/10.1088/0004-6256/145/4/101}.

@@

\bibitem[\citeproctext]{ref-navarroStructureColdDark1996}
Navarro, Julio F., Carlos S. Frenk, and Simon D. M. White. 1996. {``The
{Structure} of {Cold Dark Matter Halos}.''} \emph{The Astrophysical
Journal} 462 (May): 563. \url{https://doi.org/10.1086/177173}.
